---
title: "Barcoding primers for pooling multiple individuals"
author: "Meinolf Ottensmann"
bibliography: refs.bib
output:
    pdf_document: default
includes:
    in_header: \\usepackage[utf8]{inputnc}
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,tidy = TRUE,
                      highlight = TRUE,comment = ">", strip.white = TRUE, collapse = TRUE,
                      warning = FALSE, message = FALSE, fig.align = "center")
options(width = 50)
```

## Preface

This document provides details on barcoding individual samples using indexed primer pairs. Details on the development of target-specific primer pairs are given in files **01_Primer_Design**.

```{r, results = 'hide', eval= FALSE}
library(plyr)
library(dplyr)
library(seqinr)
library(rex)
library(ShortRead)
```


## Assing 82 samples and 6 replicates to 11 rows of a 96-well rack 

All 82 samples (41 mother-pup pairs) are randomly assigned to a location on 96-well rack. Additionally, six replicates are included.

```{r, eval = FALSE}
# read sample names
names <- read.csv("data/mhc_sample_names.csv",header = T, sep = ";")
names <- names[,c("ID","Pair")]
names <- unique(names)
row.names(names) <- as.character(1:nrow(names))

# Pseudo-random order of samples
# Note, this order will change when execute on another computer
set.seed(99)
coins <- sample(1:nrow(names),size = nrow(names),replace = F)

# create rack locations
R <- 1:11
C <- LETTERS[1:8]
p <- character()
for (r in 1:length(R)) {
  for (c in 1:length(C)) {
    p <- c(p, paste0(C[c],as.character(R[r])))
      }
}

df <- data.frame(id = names[["ID"]][coins], 
             pair = names[["Pair"]][coins], 
             no = coins, 
             pos = p[1:82])

## Last column are replicates
replicates <- c("W8914mum","W8928pup","W8933pup","W8935mum","W8915pup","AGF11007")
coins2 <- unlist(lapply(X = replicates,FUN = function(a) which(df[["id"]] == a)))

df2 <- df[coins2,1:3]
df2[["pos"]] <- p[83:88]

df <- rbind(df,df2)
df[["age"]] <- unlist(lapply(X = df[["pair"]],FUN = function(a) substring(a, 1,1)))
df[["pair_id"]] <- unlist(lapply(X = df[["pair"]],FUN = function(a) substring(a, 2)))
```

## Add barcodes

Index sequences of the Illumina Nextera_XT kit [@Illumina.2017] are used as barcodes. 12 indices may be assigned to each of the rack´s rows, while 8 indices are used for each of the columns. Note, DRB and DQB use slightly different combinations.

```{r, eval=FALSE}
# read barcodes and assing to forward and reverse
barcodes <- read.csv2("data/mhc_barcodes.csv",skip = 1)
f <- barcodes[barcodes[["primer"]] == "Forward",1:2]
r <- barcodes[barcodes[["primer"]] == "Reverse",1:2]

dqb <- 
  data.frame(Index_F = rep(f$Index[1:11],each = 8),
             Index_R = rep(r$Index, 11),
             Sequence_F = rep(f$Sequence[1:11], each = 8), 
             Sequence_R = rep(r$Sequence, 11))

drb <-
  data.frame(Index_F = rep(f$Index[c(1,2,3,4,5,6,7,12,9,10,11)], each = 8), 
             Index_R = rep(r$Index, 11),
             Sequence_F = rep(f$Sequence[c(1,2,3,4,5,6,7,12,9,10,11)], each = 8), 
             Sequence_R = rep(r$Sequence, 11))

df <- df[,c("id","age","pair_id","pos")]

dqb <- cbind(df, dqb)
drb <- cbind(df, drb)

write.csv(dqb, file = "documents/dqb_table.csv", row.names = F)
write.csv(dqb, file = "documents/dqa_table.csv", row.names = F)
write.csv(dqb, file = "documents/dra_table.csv", row.names = F)
write.csv(drb, file = "documents/drb_table.csv", row.names = F)
```

***
```{r}
sessionInfo()
```
***

## References