---
title: "Process MHC II DRB"
author: "Meinolf Ottensmann"
bibliography: refs.bib
output:
    pdf_document: default
includes:
    in_header: \\usepackage[utf8]{inputnc}
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,tidy = FALSE,
                      highlight = TRUE,comment = ">",strip.white = TRUE,collapse = TRUE,
                      warning = FALSE, message = FALSE, fig.align = "center")
options(width = 50)
```

## Preface

All processing steps for analysing the Illumina MiSeq reads are outlined within this document. Most steps are based on several `python` scripts that may be executed using the command line (e.g. 'Bash on Ubuntu in Windows'), while `R` functions can be called within this document. 

```{r}
library(magrittr)
source("R/clustering_functions.R")
```

## Prepare metafile

The first step of the processing uses the [Qiime](http://qiime.org/scripts/extract_barcodes.html) script `extract_barcodes.py` to reorient froward and reverse reads and trimming barcodes form the sequences. In order to do so, a metafile is needed that gives the forward and reverse primer sequences respectively which allow to recognise the orientation of a read. 

```{r, eval=FALSE}
## R
## read barcodes
barcodes <-
  read.csv("data/mhc_barcodes.csv", header = T, sep = ";", skip = 1)

## format left barcode
index_l <-
  barcodes[barcodes[["primer"]] == "Forward",c("Index","Sequence")]
names(index_l) <- c("index_L","BC_L")

## format right barcode
index_r <-
  barcodes[barcodes[["primer"]] == "Reverse",c("Index","Sequence")]
names(index_r) <- c("index_R","BC_R")

## read list of samples and join with barcodes
drb <- 
  read.csv("documents/drb-samples-miseq.csv", header = T, sep = "\t") %>%
  dplyr::left_join(., index_l, by = "index_L" ) %>%
  dplyr::left_join(., index_r, by = "index_R")

## create metafile to erxtract barcodes using QIIME
drb_qiime <- data.frame(
  SampleID =
    paste0(drb[["sample"]],".",drb[["position"]]),
  BarcodeSequence =
    paste0(drb[["BC_L"]],drb[["BC_R"]]),
  LinkerPrimerSequence = 
    read.table("documents/mhc_primer_sequences.csv", header = T, sep = ",")[2,2] %>%
    as.character(),
  ReversePrimer =
    read.table("documents/mhc_primer_sequences.csv", header = T, sep = ",")[2,3] %>%
    as.character(),
  Description = NA)

drb_qiime$SampleID <-
  stringr::str_replace(drb_qiime$SampleID, "/",".")

## Manually add '#' before each sample in the created file
write.table(x = drb_qiime,
            file = "miseq_reads/DRB-Pool/drb_barcodes.txt",
            sep = "\t",
            row.names = F,
            quote = F)
```
\newpage

## Start processing MiSeq reads

```bash
# bash
#######################################################################################
# Prerequisites:                                                                      #
#                                                                                     #
# 'mhc_cluster' needs to be downloded from https://github.com/mottensmann/mhc_cluster #      
# and added to the 'PATH' variable.                                                   #
# Vsearch v.2.44, hmmer-3.1b2, usearch10, muscle3.8.31 are expected to be in 'PATH'   #
#                                                                                     #
# Folder that are missing in the PATH can be added using the following notation       #
# where '~' gives the path starting from the root.                                    #
# Example: export PATH="~/mhc_cluster:$PATH"                                          #                 # Additionally QIIME is required (see http://qiime.org/install/install.html)          #
# Here, QIIME is assumed to be part of the conda environment                          #
# Raw reads may be downloaded as zip archieves and saved in a subfolder raw_reads     #
#######################################################################################

## Start conda
## -----------
source activate qiime191conda

## Set directory
## -------------
cd miseq_reads/DQB-Pool/

## MUSCLE alignment of dqb sequences
## ---------------------------------
muscle -in hmm/seal_dqb.fasta -out hmm/seal_dqb.afa ;

## create hidden markov model
## --------------------------
hmmbuild hmm/seal_dqb.hmm hmm/seal_dqb.afa ;

## create auxilariy files for hmmscan
## ----------------------------------
hmmpress hmm/seal_dqb.hmm ;

## copy to folder lib in repository mhc_cluster
## --------------------------------------------
cd hmm/
find -name "seal_dqb*" -print -exec cp {} ~/mhc_cluster/lib/ \;
cd ..

## Unzip raw reads
## ---------------
cd raw_reads
find . -iname "*.gz" -exec gunzip {} \;
cd ..

## Extract and strip barcodes from reads 
## --------------------------------------
extract_barcodes.py --input_type barcode_paired_end 
                    -f raw_reads/reads1.fastq -r raw_reads/reads2.fastq 
                    --bc1_len 8 --bc2_len 8 -m dqb_barcodes.txt 
                    -a -o parsed_barcodes/ ;
                    
## Filter reads and truncate to 230 bp
## -----------------------------------
vsearch --fastq_filter parsed_barcodes/reads1.fastq 
        -fastqout parsed_barcodes/reads1_qual_filt.fastq 
        -fastq_maxns 0 -fastq_maxee 2 -fastq_trunclen 230
```

```{r, eval = FALSE}
## R
## Filter merged reads 
## -------------------
source("R/call_filter_forward_reads_drb.R")
```

```bash
# bash
## Truncate filtered to get only exon sequences
## --------------------------------------------
fastx_trimmer -i parsed_barcodes/reads1_qual_filt_cont_filt.fastq 
              -o parsed_barcodes/reads1_cluster_input.fastq -f 31 &
              
## Run MHC clustering pipeline
## ---------------------------
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/drb 
                -hmm seal_drb.hmm 
                -alpha '2.0' 
                
## Explore sensitivity to parameter alpha
## --------------------------------------
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/alpha_exploration/drb  
                 -alpha '0.0' -hmm seal_drb.hmm ;
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/alpha_exploration/drb  
                 -alpha '0.5' -hmm seal_drb.hmm ;
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/alpha_exploration/drb  
                 -alpha '1.0' -hmm seal_drb.hmm ;
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/alpha_exploration/drb  
                 -alpha '1.5' -hmm seal_drb.hmm ;
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/alpha_exploration/drb  
                 -alpha '2.0' -hmm seal_drb.hmm ;
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/alpha_exploration/drb  
                 -alpha '2.5' -hmm seal_drb.hmm ;
cluster_mhc2.py -f parsed_barcodes/reads1_cluster_input.fastq 
                -o clustered_reads/alpha_exploration/drb  
                 -alpha '3.0' -hmm seal_drb.hmm ;
```

\newpage

```{r, eval = FALSE}
# R
## split reads in amplicons
demultiplex_fastq(
reads = "miseq_reads/DRB-Pool/parsed_barcodes/reads1_cluster_input.fastq",
out = "miseq_reads/DRB-Pool/demultiplexed",
outname = "read1.fastq")
```

```bash
# bash
## Apply to samples individually
## -----------------------------
cd /demultiplexed/
find . -type d | while read d; do
   (cd $d/
   cluster_mhc2.py -f read1.fastq -o temp  -hmm seal_drb.hmm -alpha '0.0' ;
   cluster_mhc2.py -f read1.fastq -o temp  -hmm seal_drb.hmm -alpha '0.5' ;
   cluster_mhc2.py -f read1.fastq -o temp  -hmm seal_drb.hmm -alpha '1.0' ;
   cluster_mhc2.py -f read1.fastq -o temp  -hmm seal_drb.hmm -alpha '1.5' ;
   cluster_mhc2.py -f read1.fastq -o temp  -hmm seal_drb.hmm -alpha '2.0' ;
   cluster_mhc2.py -f read1.fastq -o temp  -hmm seal_drb.hmm -alpha '2.5' ;
   cluster_mhc2.py -f read1.fastq -o temp  -hmm seal_drb.hmm -alpha '3.0' ;
   )  
done
cd ../
```

```{r, eval = FALSE}
## R
## Remove temporary created files
remove_mhc_cluster_files(
  parentfolder = "miseq_reads/DRB-Pool/demultiplexed/",
  fastq = "read1.fastq")

## Pool alleles of all individually processed amplicons
filen <- c("temp_pct_1.0_a_0.0_ee_1.0.fixed.otus.fa",
           "temp_pct_1.0_a_0.5_ee_1.0.fixed.otus.fa",
           "temp_pct_1.0_a_1.0_ee_1.0.fixed.otus.fa",
           "temp_pct_1.0_a_1.5_ee_1.0.fixed.otus.fa",
           "temp_pct_1.0_a_2.0_ee_1.0.fixed.otus.fa",
           "temp_pct_1.0_a_2.5_ee_1.0.fixed.otus.fa",
           "temp_pct_1.0_a_3.0_ee_1.0.fixed.otus.fa")
lapply(filen, function(x) {
  pool_zotus(parentfolder = "miseq_reads/DRB-Pool/demultiplexed/",
             filen = x)})

```

```bash
# bash
## Cluster using based on individually identified Zotus
## ----------------------------------------------------
cluster_against_db.py -f parsed_barcodes/reads1_cluster_input.fastq 
   -ref demultiplexed/temp_pct_1.0_a_0.0_ee_1.0.fixed.otus.fa_pooled_zotus.fa 
   -o reference_based/a_0.0 ;
cluster_against_db.py -f parsed_barcodes/reads1_cluster_input.fastq 
   -ref demultiplexed/temp_pct_1.0_a_0.5_ee_1.0.fixed.otus.fa_pooled_zotus.fa 
   -o reference_based/a_0.5 ;
cluster_against_db.py -f parsed_barcodes/reads1_cluster_input.fastq 
   -ref demultiplexed/temp_pct_1.0_a_1.0_ee_1.0.fixed.otus.fa_pooled_zotus.fa 
   -o reference_based/a_1.0 ;
cluster_against_db.py -f parsed_barcodes/reads1_cluster_input.fastq 
   -ref demultiplexed/temp_pct_1.0_a_1.5_ee_1.0.fixed.otus.fa_pooled_zotus.fa 
   -o reference_based/a_1.5  ;
cluster_against_db.py -f parsed_barcodes/reads1_cluster_input.fastq 
   -ref demultiplexed/temp_pct_1.0_a_2.0_ee_1.0.fixed.otus.fa_pooled_zotus.fa 
   -o reference_based/a_2.0 ;
cluster_against_db.py -f parsed_barcodes/reads1_cluster_input.fastq 
   -ref demultiplexed/temp_pct_1.0_a_2.5_ee_1.0.fixed.otus.fa_pooled_zotus.fa 
   -o reference_based/a_2.5 ;
cluster_against_db.py -f parsed_barcodes/reads1_cluster_input.fastq 
   -ref demultiplexed/temp_pct_1.0_a_3.0_ee_1.0.fixed.otus.fa_pooled_zotus.fa 
   -o reference_based/a_3.0 ;
```

\newpage

***
```{r}
sessionInfo()
```
***