---
title: "Analysis MHC II DQB"
author: "Meinolf Ottensmann"
bibliography: refs.bib
output:
    pdf_document: default
includes:
    in_header: \\usepackage[utf8]{inputnc}
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(fig.width = 5, fig.height = 5,tidy = FALSE,
                      highlight = TRUE,comment = ">",strip.white = TRUE,collapse = TRUE,
                      warning = FALSE, message = FALSE, fig.align = "center")
options(width = 70)
```

```{r, results='hide'}
library(ggplot2)
library(vegan)
library(magrittr)
library(ape)
library(phangorn)
library(ShortRead)
library(pegas)

source("R/clustering_functions.R")
source("R/fasta_fastq_functions.R")
source("R/genotyping_functions.R")
source("R/summary_stats.R")
```

## Explore the influcence of choosen alpha values on clustering results

The impact of the parameter alpha on clustering results was systematically investigated using individual amplicons and pooled amplicons respectively. This anlysis allows to determine how robust allele classification are given the choosen alpha value. For alpha = 0, only two alleles are retained after clustering. Therefore, allele detection using the DOC methods fails (i.e. no inflection point can be calculated). Respective datasets are not considered for the analysis


```{r}
## select datasets
fname <- c(
  "miseq_reads/DQB-Pool/reference_based/a_0.5_pct_1.0",
  "miseq_reads/DQB-Pool/clustered_reads/alpha_exploration/dqb_pct_1.0_a_0.5_ee_1.0",
  "miseq_reads/DQB-Pool/reference_based/a_1.0_pct_1.0",
  "miseq_reads/DQB-Pool/clustered_reads/alpha_exploration/dqb_pct_1.0_a_1.0_ee_1.0",
  "miseq_reads/DQB-Pool/reference_based/a_1.5_pct_1.0",
  "miseq_reads/DQB-Pool/clustered_reads/alpha_exploration/dqb_pct_1.0_a_1.5_ee_1.0",
  "miseq_reads/DQB-Pool/reference_based/a_2.0_pct_1.0",
  "miseq_reads/DQB-Pool/clustered_reads/alpha_exploration/dqb_pct_1.0_a_2.0_ee_1.0",
  "miseq_reads/DQB-Pool/reference_based/a_2.5_pct_1.0",
  "miseq_reads/DQB-Pool/clustered_reads/alpha_exploration/dqb_pct_1.0_a_2.5_ee_1.0",
  "miseq_reads/DQB-Pool/reference_based/a_3.0_pct_1.0",
  "miseq_reads/DQB-Pool/clustered_reads/alpha_exploration/dqb_pct_1.0_a_3.0_ee_1.0")

## genotype in order to classify allele status
out <-
  lapply(fname,
         run_genotyping,
         locus = "dqb",
         gain = 0.05,
         doc_min = 40,
         depth_min = 0.7)

names(out) <-
  paste0("Alpha", rep(seq(0.5,3,0.5), each = 2))

## push results in data frame
df <- lapply(out, function(x) {
  reshape2::melt(x$zotu_summary, )
  }) %>%
  do.call("rbind",.)

## format df
df$alpha <- 
  rep(seq(0.5,3,0.5), each = 6)
df$group <- 
  rep(rep(c("Individually", "Pooled"), each = 3), 6)

## make levels look nicer
levels(df$variable)[levels(df$variable) == "putatitive_artefact"] <-
  "Putatitive Artefact"
levels(df$variable)[levels(df$variable) == "putatitive_allele"] <- 
  "Putatitive Allele"

# sort factors
df$variable <-
  factor(df$variable, levels = c("Artefact", "Putatitive Artefact", "Putatitive Allele"))

## make plot
plot1_alpha <-
ggplot(df, aes(x = alpha, y = value, fill = variable)) +
  geom_col() +
  theme_classic(base_size = 14) +
  facet_grid(~group) +
  ylab("Total number of variants") +
  xlab(expression(alpha)) +
  guides(fill = guide_legend(title = "Classification")) +
  scale_y_continuous(expand = c(0,1),
                     breaks = seq(0,50,5)) +
  scale_x_continuous(expand = c(0.05,0),
                     breaks = seq(0.5,3,0.5))
plot1_alpha
```


## Read and process clustering results

```{r}
## choose data obtained from pooled sequences at alpha = 2.0
fname <- "miseq_reads/DQB-Pool/clustered_reads/alpha_exploration/dqb_pct_1.0_a_2.0_ee_1.0"
dqb_data <- process_otus(fname, locus = 'dqb')
```

## Correlation between reads counts

```{r}
## relate filtered read depth to total sequence number
lm_fit <- with(dqb_data$bc_counts, lm(Filtered_total~Raw_total))

## correlation 
with(dqb_data$bc_counts, cor.test(Filtered_total, Raw_total))

## check barcode quality among amplicons
plot2_barcode_quality <- 
ggplot(dqb_data$bc_counts, aes(x = Raw_total/1000, y = Filtered_total/1000)) +
  geom_point() +
  geom_abline(intercept = lm_fit$coefficients[[1]]/1000,
              slope = lm_fit$coefficients[[2]],
              linetype = "dashed",
              col = "blue") +
  geom_abline(intercept = 0,
              slope = 1) +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5),
        aspect.ratio = 1) +
  xlab("Total reads") +
  ylab("High quality reads") +
  scale_x_continuous(
    expand = c(0,0),
    limits = c(0, floor(max(dqb_data$bc_counts$Raw_total)/1000)),
    breaks = seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),
    labels = paste0(seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),"k"))  +
  scale_y_continuous(
    expand = c(0,0),
    limits = c(0, floor(max(dqb_data$bc_counts$Raw_total)/1000)),
    breaks = seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),
    labels = paste0(seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),"k")) 
plot2_barcode_quality

## relate mapped read depth to total sequence number
lm_fit <- with(dqb_data$bc_counts, lm(mapped~Raw_total))

## correlation 
with(dqb_data$bc_counts, cor.test(mapped, Raw_total))

plot3_barcode_mapping <- 
ggplot(dqb_data$bc_counts, aes(x = Raw_total/1000, y = mapped/1000)) +
  geom_point() +
  geom_abline(intercept = lm_fit$coefficients[[1]]/1000,
              slope = lm_fit$coefficients[[2]],
              linetype = "dashed",
              col = "blue") +
  geom_abline(intercept = 0,
              slope = 1) +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5),
        aspect.ratio = 1) +
  xlab("Total reads") +
  ylab("Mapped reads") +
  scale_x_continuous(
    expand = c(0,0),
    limits = c(0, floor(max(dqb_data$bc_counts$Raw_total)/1000)),
    breaks = seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),
    labels = paste0(seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),"k")) +
  scale_y_continuous(
    expand = c(0,0),
    limits = c(0, floor(max(dqb_data$bc_counts$Raw_total)/1000)),
    breaks = seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),
    labels = paste0(seq(0,floor(max(dqb_data$bc_counts$Raw_total)/1000),10),"k"))
plot3_barcode_mapping
```

The read depth among individuals reveals some systematic and expected matches including a tight correlation among (i) total number of reads, (ii) number of filtered reads and (iii) reads assigned to any of the alleles. However, the depth for a subset of individuals is low and a fewe samples show below average mapping success to alleles. 

## Assing genotypes to individuals

For almost all individuals sequences are assigned to each of the clustered alleles. This is expected for the followign reasons:
 
* Cross-talk/ tag switching during sequencing
* Spurios reads caused by sequencing and PCR error
* Cross-amplicon contamination

For these reasons, genotyping requires to separate true alleles from artefacts. True alleles are expected to be more common than any of the spurious reads present in a given amplicon. This motivates to use the degree of change (DOC) approach suggested by [-@Lighten.2014] that determines inflection points in the cumulative sequencing depth. In contast to commonly used methods that are based on the sequencing depth [@Babik.2009, @Galan.2010], this approach does not rely on any arbitrary cut-off value but directly tests the main genotyping assumption outlined above. 

```{r}
## Calculate cumulative sequencing depth for every amplicon.
## Then, call alleles and assign quality class.
genotypes_list <- apply(dqb_data$otu_tab, 2,
                        get_genotypes,
                        names = rownames(dqb_data$otu_tab),
                        gain = 0.05,
                        doc_min = 40,
                        depth_min = 0.75) 

## get cumulative sums
genotypes_df <- do.call("rbind", lapply(genotypes_list, function(x) x[["coord"]])) %>%
  subset(., quality == "High")
genotypes_df$group <- as.factor(genotypes_df$group)

## Calculate mean relative cumulative sequencing depth
df <- summary_stats(data = genotypes_df, measurevar = "y", groupvars = c("x", "group"))

qual <- lapply(genotypes_list, function(x) x[["df"]][[6]])
summary(unlist(qual))

# add point x = 0, y = 0 for visualisation
df_head <- matrix(0, 
                   nrow = length(levels(df$group)),
                   ncol = ncol(df)) %>%
  as.data.frame() %>%
  set_colnames(., colnames(df))
df_head$group <- 1:nrow(df_head)
df <- rbind(df_head, df)

## Get mean depth for all allele number configurations
df_dashes <- data.frame(x = 1:4,
                        group = as.character(1:4),
                        y = c(df$y[df$group == 1 & df$x == 1],
                              df$y[df$group == 2 & df$x == 2],
                              df$y[df$group == 3 & df$x == 3],
                              df$y[df$group == 4 & df$x == 4]))

plot4_cumul_depth <- ggplot(df, aes(x = x, y = y, col = group)) +
  geom_point(size = 1.75) +
  geom_line(size = 1) +
  geom_segment(data = df_dashes, 
               aes(xend = x, yend = 0),
               linetype = "dotted",
               size = 1) +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "none") +
  xlab("Number of alleles") +
  ylab("Cumulative sequencing depth [%]") +
  scale_y_continuous(expand = c(0,0),
                     breaks = seq(0,100,10)) +
  scale_x_continuous(expand = c(0,0),
                     breaks = 0:8) +
  scale_color_brewer(palette = "Set2")
plot4_cumul_depth

  ## export plots for each sample
  genotypes <- apply(dqb_data$otu_tab, 2,
                        get_genotypes,names = rownames(dqb_data$otu_tab),
                     plot = T)
  for (i in 1:length(genotypes)) {
    genotypes[[i]] <- genotypes[[i]] + ggtitle(names(genotypes)[i])
  }

pdf("miseq_reads/DQB-Pool/figures/genotypes_DOC.pdf")
for (i in  1:length(genotypes)) print(genotypes[[i]])
dev.off()

## Summarise by amplicon sequencing depth

sequencing_depth <- do.call("rbind", lapply(genotypes_list, function(x) x[["df"]])) %>%
  as.data.frame()

## amplicon quality scores
summary(sequencing_depth$quality)

plot5_amplicon_quality <- ggplot(sequencing_depth, aes(x = quality, y = total_depth)) +
  geom_boxplot() +
  theme_classic(base_size = 14) +
  xlab("Amplicon quality") +
  ylab("Log total sequencing depth") +
  scale_y_log10()
plot5_amplicon_quality

with(sequencing_depth, wilcox.test(total_depth~quality))

allele_num_df <- lapply(genotypes_list, function(x) x[["df"]]) %>%
  do.call("rbind",.) %>%
  subset(., quality == "High")
allele_num_df$row <- rownames(allele_num_df)

plot6_allele_num_depth <- 
  ggplot(allele_num_df, 
         aes(x = as.factor(n_alleles), y = total_depth,
            fill = as.factor(n_alleles))) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Paired") +
  theme_classic(base_size = 14) +
  theme(legend.position = "none") +
  xlab("Number of alleles") +
  ylab("Total sequencing depth") 
plot6_allele_num_depth

with(allele_num_df, kruskal.test(total_depth~n_alleles))
with(allele_num_df, pgirmess::kruskalmc(total_depth~n_alleles))
```


The above boxplot shows that there is no signifcant linear trend of increasing number of allels with respect to the total sequencing depth. However, theres is a clear different in accordance with one or two amplified loci.

## Exploring potential bias by variation in sequencing depth

The function `vegan::rarefy` gives the expected species richness in random subsamples of size sample and therefore allows to test if variation in sequencing depth could have an effect on the detection of alleles.

```{r}
## get sequence counts of high quality amplicons
otu_table <- 
  dqb_data$otu_tab[rownames(allele_num_df[allele_num_df$quality == "High",])] %>%
  t()

## define sample sizes. Consider 0.5 of minimum depth as starting point
size <- seq(from = floor(min(rowSums(otu_table))*0.5),
                to = min(rowSums(otu_table)),
                by = 50)

## Conduct rarefaction analysis for each sample size
df <- data.frame(
  sample = rep(rownames(otu_table), length(size)),
  size = rep(size, each = length(rownames(otu_table))),
  y = unlist(lapply(size, function(x) {
    rarefaction(m = otu_table, 
                n = x, 
                gain = 0.05,
                doc_min = 40,
                depth_min = 0.75)})))

plot7_rarefy <- ggplot(df, aes(x = size, y = y, grp = sample)) + 
  geom_jitter(col = "grey50", alpha = 0.2) +
  geom_line(size = 1.5, col = "darkorange") +
  theme_classic(base_size = 14) +
  scale_x_continuous(
    expand = c(0,0),
    breaks = floor(seq(from = min(size),
                 to = plyr::round_any(max(size), 1000, f = ceiling),
                 length.out = 10))) +
  scale_y_continuous(breaks = seq(0, max(df$y))) +
  xlab("Total sequencing depth") +
  ylab("Number of detected alleles")
plot7_rarefy
```
 

```{r}
## Call alleles for the final dataset
called_alleles <- apply(otu_table, 1, function(x) {
  out <- get_genotypes(x,
                       gain = 0.05,
                       doc_min = 40,
                       depth_min = 0.75)
  out[["alleles"]]
})
save(called_alleles, file = "miseq_reads/DQB-Pool/RData/called_alleles.RData")

## Rarefy to minimum depth 
df <- data.frame(
  obs = unlist(lapply(called_alleles, length)),
  rarefied = rarefaction(m = otu_table,
                         n = min(rowSums(otu_table)),
                         gain = 0.05,
                         doc_min = 40,
                         depth_min = 0.75))

## summarise
df_sum <- summary_stats(df, measurevar = "obs", groupvars = c("rarefied"))

plot8_rarefied_depth <-
ggplot(df_sum, aes(x = obs, y = rarefied, size = N)) +
  geom_point(col = "blue") +
  scale_size(range = c(4, 6), 
             limits = range(df_sum$N),
             breaks = unique(df_sum$N),
             name = "N") +
  geom_abline(intercept = 0, slope = 1,
              linetype = "dashed") +
  theme_classic(base_size = 14) +
  xlab("Observed number of alleles") +
  ylab("Rarefied number of alleles") +
  scale_x_continuous(breaks = seq(0,10,1)) +
  scale_y_continuous(breaks = seq(0,10,1))
plot8_rarefied_depth
```

The above graphs show that for all retained samples with a minimum sequencing depth of `r min(rowSums(otu_table))` genotypes are obustly assigned. 

## Show Barplot of amplicon sequencing depth and assigned quality

```{r}
df <- dqb_data$bc_counts[,c("pos", "mapped")]

df$grp <- ifelse(df$pos %in% names(called_alleles), 'High', 'Low')
df$pos <- factor(df$pos, levels =  as.character(df$pos))

plot9_retained_samples <- ggplot(df, aes(x = pos, y = mapped, fill = grp)) +
  geom_bar(stat = "identity", colour = "black") +
  theme_classic(base_size = 14) +
  xlab("Sample position") +
  ylab("Sequencing depth") +
  scale_fill_manual(values = c("#377EB8", "#E41A1C")) +
  theme(axis.text.x =  element_blank()) +
  guides(fill = guide_legend(title = "Amplicon quality")) 
plot9_retained_samples
summary(as.factor(df$grp))
```

## Distribution of alleles

```{r}
df <- unlist(called_alleles) %>%
  as.factor() %>%
  summary()

allele_order <- lapply(names(df), function(x) strsplit(x, split = "Zotu")[[1]][2]) %>%
  unlist()  %>%
  as.numeric() %>%
  order(., decreasing = T)

df <- data.frame(x = names(df), y = df)
df$x <- unlist(lapply(df$x, function(x) {
  stringr::str_replace(x, "Zotu", "ArGa-DQB*")
  }))
df$x <- factor(df$x, levels = df$x[allele_order])
df$z <- ifelse(df$y == 1,  "Potential artefact", "Putatitive allele")


plot10_allele_freq <- ggplot(df, aes(x = x, y = y, fill = z)) +
  geom_bar(stat = "identity", colour = "black") +
  theme_classic(base_size = 14) +
  xlab("Allele") +
  ylab("Frequency") +
  scale_y_continuous(expand = c(0,0),
                     breaks = seq(0, max(df$y), by = 3)) +
  scale_fill_brewer(palette = "Set1") +
  guides(fill = guide_legend(title = "")) +
  coord_flip()
plot10_allele_freq
```


```{r, eval = FALSE, echo = FALSE}
a_2.0_pooled <- 
  readFasta("miseq_reads/DQB-Pool/clustered_reads/dqb_pct_1.0_a_2.0_ee_1.0.fixed.otus.fa") %>%
  sread() %>%
  as.character()

a_2.0_ind <- 
readFasta("miseq_reads/DQB-Pool/merged_reads/demultiplexed/dqb_pct_1.0_a_2.0_ee_1.0.fixed.otus.fa_pooled_zotus.fa") %>%
  sread() %>%
  as.character()

comb <- c(a_2.0_pooled, a_2.0_ind) %>%
  unique()
length(comb)
}
```


## Replace rack positions by id names 

```{r}
## get mapping file
load("miseq_reads/DQB-Pool/RData/dqb_mapping_file.RData")
dqb_mapping_file <- 
  subset(dqb_mapping_file, as.character(pos) %in% names(called_alleles))
## get factors
factors <-
  read.table(file = "data/factors.txt") %>%
  subset(.,rownames(.) %in% dqb_mapping_file$Pair)
## set pair id as rowname
factors$Pair <- rownames(factors)

## merge data sheets
dqb_mapping_file <- dplyr::left_join(dqb_mapping_file, factors, by = "Pair")
dqb_mapping_file$Pair <- paste0(dqb_mapping_file$Pair,"_" , 1:nrow(dqb_mapping_file))

## substitute rack location by pair sample identifier
called_alleles_renamed <- called_alleles 
names(called_alleles_renamed) <- 
  dqb_mapping_file$Pair[match(names(called_alleles),dqb_mapping_file$pos)]

## get sample names
samples <- lapply(names(called_alleles_renamed), function(x) strsplit(x, "_")[[1]][1])

## find replicated samples
replicated <- samples[duplicated(samples)]

## randomly discard one sample of a pair of repliacates
set.seed(999)
remove <- numeric()
for (i in 1:length(replicated)) remove[i] <- sample(which(samples == replicated[[i]]), 1)

genotypes_dqb <- called_alleles_renamed[-remove]

## remove name extensions
names(genotypes_dqb) <- 
  lapply(names(genotypes_dqb), function(x) strsplit(x, "_")[[1]][1]) %>%
  unlist()

save(genotypes_dqb, file = "miseq_reads/DQB-Pool/RData/genotypes_dqb.RData")
```


## Median Joining network

```{r, eval = FALSE}
x <- fasta2mat(fasta = "miseq_reads/DQB-Pool/clustered_reads/dqb_pct_1.0_a_2.0_ee_1.0.fixed.otus.fa")

# Median-Joining Network
mjn1 <- mjn(x = x, prefix = "", epsilon = 0)

pdf(file = "miseq_reads/DQB-Pool/plots/mjn.pdf",
    width = 14,
    height = 14)
plot(mjn1, col.link = "grey80", pie = matrix(1, nrow = length(attr(mjn1, "labels"))), font = 1, cex = 0.8, scale.ratio = 1) 
dev.off()
```

## Mapping alleles to Genome and Transcriptome

Alleles map all to Contig 48. The top hit represents the expected location of the DQB locus, wheras the second best hit shows the DRB locus. Note, that only the top hit reaches a full-lenght alignment. Arga-DQB*10 is identical to the consensus sequence of the genome. 

```bash
blastn 
  -db linux/db/arc_gaz_genome_db 
  -outfmt 6 
  -num_threads 8 
  -evalue 1e-8 
  -word_size 7 
  -query miseq_reads/DQB-Pool/clustered_reads/dqb_pct_1.0_a_2.0.fixed.otus.fa 
  -out  miseq_reads/DQB-Pool/fasta/dqb2_arc_gaz_genome.fasta

blastn 
  -db linux/db/arc_gaz_transcriptome_db 
  -outfmt 6 
  -num_threads 8 
  -evalue 1e-8 
  -word_size 7 
  -query miseq_reads/DQB-Pool/clustered_reads/dqb_pct_1.0_a_2.0.fixed.otus.fa 
  -out  miseq_reads/DQB-Pool/fasta/dqb2_arc_gaz_transcriptome.fasta
```

```{r}
read.table("miseq_reads/DQB-Pool/fasta/dqb2_arc_gaz_genome.fasta")[28:34,c(1:4,9,10)] %>%
  set_colnames(., value = c("Allele", "Contig","Identity","Alignment", "Start", "End")) %>%
  head()
```

Alleles map to a single region of the assembled transcriptome.

```{r}
read.table("miseq_reads/DQB-Pool/fasta/dqb2_arc_gaz_transcriptome.fasta")[,c(1:4,9,10)] %>%
  set_colnames(., value = c("Allele", "Contig","Identity","Alignment", "Start", "End")) %>%
  head()
```

## Check transcriptome reads for expression of alleles

In order to check for evidence of gene expression for the newly characterised, alleles are mapped to raw
transcriptome reads published by [@Hoffman.2013]. These sequences are available as an archieve on Genbank under accession number SRA064103. Reads are available as separate files that will be firt downloaded and then merge into a single file that is ready to use for blasting. Here the maximum number of 45 alleles retained after clustering amplicons individually with (`r expression(alpha)` = 3.0) all 45 alleles that include 18 putative alleles, 3 putatitive artefacts and 24 sequences without support.

```bash
## bash
cd DQB

## Download 454 raw reads
## ----------------------
vdb-dump -f tab -C READ SRR646623 | awk '{print ">" "heart." NR "\n" $0}' 
         > heart.SRA064103.fasta &
vdb-dump -f tab -C READ SRR646624 | awk '{print ">" "intestine." NR "\n" $0}' 
         > intestine.SRA064103.fasta & 
vdb-dump -f tab -C READ SRR646625 | awk '{print ">" "kidney." NR "\n" $0}' 
         > kidney.SRA064103.fasta &
vdb-dump -f tab -C READ SRR646626 | awk '{print ">" "lung." NR "\n" $0}' 
         > lung.SRA064103.fasta &
vdb-dump -f tab -C READ SRR646627 | awk '{print ">" "spleen." NR "\n" $0}' 
         > spleen.SRA064103.fasta &
vdb-dump -f tab -C READ SRR646628 | awk '{print ">" "testis." NR "\n" $0}' 
         > testis.SRA064103.fasta &

## Merge files
## -----------
cat *.SRA064103.fasta > transcriptome_reads.fasta

## Dereplicate alleles:
## --------------------
usearch10.exe -fastx_uniques arga_dqb.fasta -fastaout arga_dqb_uniques.fasta

## Remove size annotation
## ----------------------
usearch10.exe -fastx_strip_annots arga_dqb_uniques.fasta -fastaout arga_dqb_derep.fasta

## make blast database from allele sequences
## -----------------------------------------
makeblastdb -in arga_dqb_derep.fasta -dbtype nucl -out arga_dqb_db

## Blast 454 reads to alleles
## --------------------------
blastn -db arga_dqb_db -outfmt 6 -num_threads 8 -evalue 1e-8 -word_size 7 -query transcriptome_reads.fasta -out transcriptome_reads.arga_dqb.txt &
```

```{r}
blastn_output <-
  read.table("DQB/transcriptome_reads.arga_dqb.txt")[,1:6] %>%
  set_colnames(., value = c("Query", "Allele", "Similarity",
            "Length", "Mismatches", "Gaps")) %>%
  subset(., Similarity >= 95 & Mismatches == 0 & Length >= 180)  
    
seqs <- readFasta("DQB/transcriptome_reads.fasta")
seqs <- seqs[which(id(seqs) %in% blastn_output$Query)]
writeFasta(object = seqs,
           file = "DQB/transcriptome_reads.arga_dqb.fasta")
write.table(x = blastn_output, 
            file = "DQB/transcriptome_reads.arga_dqb.hits.metafile.txt",
            row.names = F)
```

```{r}
expressed <- read.csv("DQB/expressed_arga_dqb.csv")
expressed$Allele <- factor(expressed$Allele, levels = paste0("ArGa-DQB*", c(11,5,4,2,1)))

plot11_gene_expression <-
ggplot(expressed, aes(x = Allele, y = 1, fill = Organ)) +
  geom_bar(stat = "identity") +
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 14),
        axis.text.y = element_text(size = 12),
        axis.text.x =  element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.x = element_blank()) +
  xlab("") +
  ylab("") +
  coord_flip() +
  scale_fill_brewer(palette = "Set2")
plot11_gene_expression
```


### Estimating differential amplification efficiencies across alleles

In a study on MHC II DRB in a rodent, Sommer *et al.* [-@Sommer.2013] have shown remarkable variation in the amplfication efficiencies differing by more than magnitude. Based on the read counts per allele across the studied population, 

```{r}
otu_table_purified <- purify_otus(x = otu_table,
                                  y = called_alleles)
## number of alleles
nb.alleles <- ncol(otu_table_purified)

## efficiency prior, all are equal i.e. 1
efficiency_prior <- rep(1, nb.alleles)

## Fit relative efficiencies based on Loglikelyhood
efficiency_obs_rel <- optim(par = efficiency_prior,
                            fn = LoglikData,
                            data = otu_table_purified,
                            control = list(fnscale = -1),
                            method = "L-BFGS-B",
                            lower = rep(0.1, nb.alleles), 
                            upper = rep(6, nb.alleles))

## Standardise effiencies with respect to ArGa-DQB*1 
efficiency_obs_norm <- 
  efficiency_obs_rel$par/efficiency_obs_rel$par[which(colnames(otu_table) == "Zotu1")]

## create a data frame
df <- data.frame(allele = colnames(otu_table),
                 efficiency = efficiency_obs_norm)
df$allele <- factor(df$allele, levels = paste0("Zotu", 1:nrow(df)))

plot12_amplification_efficiency <-
ggplot(df, aes(x = allele, y = efficiency)) +
  geom_point() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        text = element_text(size = 12)) +
  xlab("Allele") + 
  ylab("Standardised amplification efficiency") +
  scale_y_continuous(breaks = seq(0,1,0.2),
                     limits = c(0,1))
plot12_amplification_efficiency
```

## Allele detection curve

```{r}
sample_alleles <- function(data, n = seq(1, length(data),2), bs = 9999) {
  x <- rep(n, each = bs)
  y <- lapply(x, function(temp) {
    get <- data[sample(x = 1:length(data),
            size = temp,
            replace = T)] %>%
  unlist() %>%
  unique() %>%
  length()
  }) 
  
  df <- data.frame(x = x,y = unlist(y))
  df$x <- as.factor(df$x)
  return(df)
}

sampled_alleles_df <- sample_alleles(data = called_alleles)
sampled_alleles_df$x <- as.factor(sampled_alleles_df$x)

df <- summary_stats(sampled_alleles_df,
                    measurevar = "y",
                    groupvars = "x",
                    conf.interval = 0.95)
df$x <- as.numeric(as.character(df$x))

plot13_allele_detection <- ggplot(df, aes(x,y)) +
  geom_point(col = "darkorange", size = 2) +
  annotate("point", x = 13, y = 4, colour = "blue") +
  annotate("text", x = 13, y = 3.5, label = "Hoelzel 1999") +
    geom_linerange(ymin = df$y - df$sd, ymax = df$y + df$sd) +
  theme_classic(base_size = 12) +
  theme(aspect.ratio = 1) +
  xlab("Sample size") +
  ylab("Number of alleles") +
  scale_x_continuous(breaks = seq(0,length(data),4)) +
  scale_y_continuous(breaks = seq(0,21,3),
                     limits = c(0,21))
plot13_allele_detection
```


```{r, echo=FALSE}
plots <- sapply(ls(pattern = "^plot"), get)
pdf(paste0("miseq_reads/DQB-Pool/figures/","Diagnostics.pdf"))
silent <- lapply(plots, print)
dev.off()

for (i in 1:length(plots)) {
  ggsave(plots[[i]], 
         filename = paste0("miseq_reads/DQB-Pool/plots/", names(plots)[i], ".tiff"),
         width = 6,
         height = 4,
         units = "in",
         dpi = 300,
         device = "tiff")
}
```



***
```{r}
sessionInfo()
```
***

## References